### 2019-01-07 arxiv-version/main.tex

\documentclass[12pt]{article}

% LaTeX packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[singlelinecheck=false]{caption}
\usepackage{fullpage}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{thmtools}
\usepackage{tikz}
\usepackage{color}
\usepackage{comment}
\usepackage{commath}
\usepackage{dsfont}

% TikZ packages
\usetikzlibrary{arrows}
\usetikzlibrary{angles}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{cd}  % commutative diagrams
\usetikzlibrary{intersections}
\usetikzlibrary{patterns}
\usetikzlibrary{shapes}
\usetikzlibrary{through}

% Math environments
\newtheorem{definition}{Definition}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{corollary}[definition]{Corollary}

% Math symbols and commands
\newcommand{\calF}{\mathcal{F}}
\newcommand{\one}{\mathds{1}}
\newcommand{\R}{\mathbb{R}}  % set of real numbers
\newcommand{\indicator}[1]{\mathbf{1}\left[#1 \right]} % indicator
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle} % inner product
%\newcommand{\norm}[1]{\left\| #1 \right\|}  % norm of a vector or a matrix
%\DeclareMathOperator*{\deg}{deg}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\Exp}{\mathbf{E}}  % expected value
\DeclareMathOperator{\diff}{d \!} % differential
\DeclareMathOperator*{\polylog}{polylog}

\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkred}{rgb}{0.7,0,0}
\definecolor{teal}{rgb}{0.3,0.8,0.8}
\definecolor{orange}{rgb}{1.0,0.5,0.0}
\definecolor{purple}{rgb}{0.8,0.0,0.8}
\definecolor{blue}{rgb}{0.0,0.0,1.0}
\newcommand{\kibitz}[2]{{\textcolor{#1}{\textsf{\footnotesize #2}}}}
\newcommand{\alina}[1]{\kibitz{darkred}{[AB: #1]}}
\newcommand{\david}[1]{\kibitz{darkgreen}{[DP: #1]}}
\newcommand{\balazs}[1]{\kibitz{purple}{[BS: #1]}}
\newcommand{\deva}[1]{\kibitz{teal}{[DT: #1]}}
\newcommand{\chenyu}[1]{\kibitz{orange}{[CW: #1]}}
\newcommand{\chicheng}[1]{\kibitz{blue}{[CZ: #1]}}

\title{Online Multiclass Linear Classification with Bandit Feedback}

\author{
Alina Beygelzimer \and
D\'avid P\'al \and
Bal\'azs Sz\"or\'enyi \and
Devanathan Thiruvenkatachari \and
Chen-Yu Wei \and
Chicheng Zhang
}

\begin{document}

\maketitle

\begin{abstract}
We study the problem of efficient online multiclass linear classification with bandit feedback, where all examples belong to one of $K$ classes and lie in the $d$-dimensional Euclidean space.
Previous works have left open the challenge of designing efficient algorithms with finite mistake bounds
when the data is linearly separable by a margin $\gamma$.
In this work, we take a first step towards this problem.
We identify two notions of linear separability:
one-versus-all (OvA) linear separability and weakly linear separability.
\begin{enumerate}
\item Under the OvA
linear separability condition, we design an efficient algorithm that achieves a near-optimal mistake bound of $O(\frac{K}{\gamma^2})$.
\item Under the more challenging weakly linear separability condition, we design an efficient algorithm with a mistake bound of $2^{\tilde{O}(\min(K \log^2 \frac 1 \gamma, \log K \sqrt{\frac 1 \gamma})) }$ \footnote{We use the notation $\tilde{O}(f(\cdot)) = O( f(\cdot) \polylog(f(\cdot)) )$.}.
Our algorithm is based on a reduction from weakly linear separability to OvA linear separability via an infinite dimensional monomial feature map, which in turn is inspired by ~\cite{Klivans-Servedio-2008}'s work on learning intersection of halfspaces.
\end{enumerate}

%We focus on the setting where there are $K$ classes, and there is a $d$-dimensional multiclass linear classifier that separates all the examples shown with a positive margin $\gamma$.
%$2^{O(\min(K \log^2 K \log^2 \frac 1 \gamma, \log K \log \log K \sqrt \frac 1 \gamma \log \frac 1\gamma))}$
%$\min( (\log K \sqrt{\frac 1 \gamma})^{O(\log K \sqrt{\frac 1 \gamma})}, K^{O(K \log K \log^2 \frac1\gamma)} )$.

\end{abstract}

\input{introduction}

\input{related-work}

\input{notions-of-linear-separability}

\input{algorithm-for-strongly-linearly-separable-data}

\input{from-linear-separability-to-strong-linear-separability}

\input{proofs-of-polynomial-approximation-theorems}

\input{nearest-neighbor-algorithm}

\input{np-hardness-of-the-weak-labeling-problem}

\input{mistake-lower-bound-for-ignorant-algorithms}

\bibliographystyle{plainnat}
\bibliography{biblio}

\appendix

\input{appendix-multiclass-perceptron}

\end{document}