### 2019-01-07 arxiv-version/mistake-lower-bound-for-ignorant-algorithms.tex

\section{Mistake lower bound for ignorant algorithms}

\begin{lemma}
For any $\gamma \in (0,\frac 1 {160})$,
there exists a set of $N$ examples $x_1, \ldots, x_N$ and vectors $w_1, \ldots, w_N$ ($N = (\Omega(\frac{1}{\gamma}))^{(d-2)/2}$), such that:
$\| x_i \|_2 \leq 1$, $\| w_i \|_2 \leq 1$ for all $i$ in $[N]$, and for all $i, j$ in $[N]$,
\[ w_i \cdot x_j \begin{cases} \geq \gamma ,& i = j, \\ \leq -\gamma ,& i \neq j. \end{cases} \]
\label{lem:embed_d_gamma}
\end{lemma}
\begin{proof}
By Lemma 6 of~\cite{Long1995}, there exists a set of $N = (\frac{1}{2\sqrt{40\gamma}})^{d-2}$ $(d-1)$-dimensional vectors $z_1, \ldots, z_N$ on the unit sphere $\mathbb{S}^{d-2}$, such that
$\theta(z_i, z_j) \geq \sqrt{40 \gamma}$. As $\cos\theta \leq 1-\theta^2/5$ for every $\theta \in [0,\pi]$,
this implies that
\[ z_i \cdot z_j \begin{cases} = 1 ,& i = j, \\ \leq 1- 8\gamma ,& i \neq j. \end{cases} \]

Define $x_i = (\frac12 z_i, \frac12)$, and $w_i = (\frac 12 z_i, -\frac 12(1-4\gamma))$ for all $i$ in $[N]$. It can be easily checked that for all $i$, $\|x_i\| \leq 1$ and $\|w_i\| \leq 1$. In addition,
\[ w_i \cdot x_j = \frac 1 4 z_i \cdot z_j - \frac {1-4\gamma} 4 \begin{cases} \geq \gamma ,& i = j, \\ \leq -\gamma ,& i \neq j. \end{cases}\]
\end{proof}