### 2019-01-09 ICML-submission/main-supplementary.tex

\documentclass{article}

\input{commands}

\begin{document}

\input{title-and-abstract}
\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}

\bibliography{biblio}
\bibliographystyle{icml2019}

% For actual submission remove everything below.
% Use appendices only in supplementray material.
\appendix
\clearpage

\end{document}