### 2019-01-09 ICML-submission/title-and-abstract.tex

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Online Multiclass Linear Classification with Bandit Feedback}

\twocolumn[
\icmltitle{Online Multiclass Linear Classification with Bandit Feedback}

\begin{icmlauthorlist}
\icmlauthor{Alina Beygelzimer}{yahoo}
\icmlauthor{D\'avid P\'al}{yahoo}
\icmlauthor{Bal\'azs Sz\"or\'enyi}{yahoo}
\icmlauthor{Devanathan Thiruvenkatachari}{nyu}
\icmlauthor{Chen-Yu Wei}{usc}
\icmlauthor{Chicheng Zhang}{microsoft}

\end{icmlauthorlist}

\icmlaffiliation{yahoo}{Yahoo Research, New York, NY, USA}
\icmlaffiliation{nyu}{New York University, New York, MA, USA}
\icmlaffiliation{usc}{University of Southern California, Los Angeles, CA, USA}
\icmlaffiliation{microsoft}{Microsoft Research, New York, NY, USA}

\icmlcorrespondingauthor{D\'avid P\'al}{davidko.pal@gmail.com}

\icmlkeywords{multi-armed bandits, contextual bandits, online classification, linear separability}

\vskip 0.3in
]

\printAffiliationsAndNotice{}

\begin{abstract}
We study the problem of efficient online multiclass linear classification with
bandit feedback, where all examples belong to one of $K$ classes and lie in the
$d$-dimensional Euclidean space. Previous works have left open the challenge of
designing efficient algorithms with finite mistake bounds when the data is
linearly separable by a margin $\gamma$. In this work, we take a first step
towards this problem. We consider two notions of linear separability: strong
linear separability and weak linear separability.

\begin{enumerate}
\item Under the strong linear separability condition, we design an efficient
algorithm that achieves a near-optimal mistake bound of
$O\left(\frac{K}{\gamma^2} \right)$.

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a infinite-dimensional feature mapping that transforms weak linear
separable examples to strong linear separable examples, which in turn is
inspired by \cite{Klivans-Servedio-2008}'s work on learning intersection of
halfspaces.
\end{enumerate}
\end{abstract}