### 2019-01-15 ICML-submission/main-all.tex
Switch to one .tex file

\documentclass{article}

\input{commands}

\begin{document}

\input{title-and-abstract}
\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}
\input{algorithm-for-strongly-linearly-separable-data}
\input{from-weak-separability-to-strong-separability}

\bibliography{biblio}
\bibliographystyle{icml2019}

% For actual submission remove everything below.
% Use appendices only in supplementray material.
\appendix
\clearpage

\input{appendix-multiclass-perceptron}
\input{appendix-proofs-of-polynomial-approximation-theorems}

\end{document}