### 2019-01-17 ICML-submission/related-work.tex
another pass over related work

\section{Related work}
\label{section:related-work}

The problem of online bandit multiclass learning was initially formulated in the
pioneering work of~\citet{Auer-Long-1999}, under the name of ``weak
reinforcement model''. They showed that if all examples agree with some
classifier from a prespecified hypothesis class $\calH$, then the optimal
mistake bound in the bandit setting can be upper bounded by the optimal mistake
bound in the full information setting, times a factor of $(2.01 + o(1))K \ln K$.
\citet{Long-2017} later improved the factor to $(1 + o(1)) K \ln K$ and showed
its near-optimality. \citet{Daniely-Helbertal-2013} extended the results to the
setting where the performance of the algorithm is measured by its regret, i.e.
the difference between the number of mistakes made by the algorithm and the
number of mistakes made by the best classifier in $\calH$ in hindsight.
We remark that
all algorithms developed in this context are computationally inefficient.

The linear classification version of this problem is initially studied
by~\citet{Kakade-Shalev-Shwartz-Tewari-2008}. They proposed two computationally
inefficient algorithms that work in the weakly linearly separable setting, one
with a mistake bound of $O(K^2 d \ln(\frac{d}{\gamma}))$, the other with a
mistake bound of $\widetilde{O}(\frac{K^2}{\gamma^2} \ln T)$. The latter result
was later improved by \citet{Daniely-Helbertal-2013}, which gives a
computationally inefficient algorithm with a mistake upper bound of
$\widetilde{O}(\frac{K}{\gamma^2})$. In addition, the authors propose the
\textsc{Banditron} algorithm, a computationally efficient algorithm that has a
$O(T^{2/3})$ regret against the multiclass hinge loss in the general setting,
and has a $O(\sqrt{T})$ mistake bound in the $\gamma$-weakly linearly separable
setting.
In contrast to mild dependencies on the time horizon for mistake bounds of
computationally inefficient algorithms,
the polynomial dependence of \textsc{Banditron}'s mistake bound on the
time horizon is undesirable for problems with a long time horizon, in the
weakly linearly separable setting. One key open
question left by~\citet{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can
design computationally efficient algorithms that achieve mistake bounds that
match or improve over those of inefficient algorithms. In this paper, we take a
step towards answering this question, showing that efficient algorithms with
mistake bounds quasipolynomial in $\frac 1 \gamma$ or quasipolynomial in $K$ can
be obtained.

%In addition, it also
% shows that under different relationships between $k$, $d$ and $\gamma$,
% the adversary can force the learner to incur $\Omega(K^2 d)$
% or $\Omega(\frac{K}{\gamma^2})$ mistakes.

%Whether one can design an efficient algorithm with a finite mistake bound that
%has no dimensionality dependence is mentioned as an open problem in
%~\cite{Kakade-Shalev-Shwartz-Tewari-2008}, where in this paper we provide a
%positive answer. $O(k^2 d \ln(\frac{d}{\gamma^2}))$

Many works consider bandit multiclass classification in the general setting,
where the examples are not necessarily separable by some linear
classifier~\cite{Abernethy-Rakhlin-2009, Wang-Jin-Valizadegan-2010,
Crammer-Gentile-2013, Hazan-Kale-2011, Beygelzimer-Orabona-Zhang-2017,
Foster-Kale-Luo-Mohri-Sridharan-2018}. In addition, \citet{Chen-Lin-Lu-2014,
Zhang-Jung-Tewari-2018} study online bandit multiclass boosting under bandit
feedback, where one can view boosting as linear classification by treating each
base hypothesis as a separate feature. However, most of these works achieve a
regret of order $\widetilde{O}(\sqrt{T})$ to $O(T^{3/4})$.
In the weakly linearly
separable setting, these algorithms at best guarantees a mistake upper bound of
$\widetilde{O}(\sqrt{T})$, which is undesirable for problems with a long time
horizon. One exception is the \textsc{Newtron} algorithm, where with appropriate
settings of parameter $\alpha$, it is able to achieve a $O(\ln T)$ regret against
the $\alpha$-logistic loss.
However, as argued in ~\citep[][Appendix D]{Beygelzimer-Orabona-Zhang-2017}, if
the setting of $\alpha$ makes the regret bound $O(\sqrt{T})$, the $\alpha$-logistic loss
of the best linear predictor would be $\Omega(T^{3/4})$. Therefore,
\textsc{Newtron} at best guarantees a mistake upper bound of
$O(\sqrt{T})$ in the weakly linearly separable setting.

%\citet{Abernethy-Rakhlin-2009} poses the open problem of whether one can design
%an efficient algorithm to get a $O(\sqrt{T})$ regret against some reasonable
%loss functions. \citet{Crammer-Gentile-2013} show that such an algorithm can be
%obtained, provided that the distribution of the label $y_t$ conditioned on
%feature vector $x_t$ satisfies certain parametric noise assumption.
%\citet{Hazan-Kale-2011} developed the \textsc{Newtron} algorithm, which has a
%regret between $O(\ln T)$ and $O(T^{2/3})$ against the multiclass logistic
%loss, where the exact order of the regret depends on the diameter of the
%benchmark class. In particular, if the diameter is $O(\ln T)$, its regret bound
%would become $O(T^{2/3})$. The \textsc{SOBA} algorithm by
%\citet{Beygelzimer-Orabona-Zhang-2017} achieves a regret of
%$\widetilde{O}(\sqrt{T})$ against the
%$\eta$-loss~\citet{Orabona-Cesa-Bianchi-Gentile-2012}. In addition, its regret
%bound does not depend sensitively on the diameter of the benchmark class.
%\citet{Foster-Kale-Luo-Mohri-Sridharan-2018} developed an algorithm that has a
%regret of $\widetilde{O}(\sqrt{T})$ against the multiclass logistic loss, where
%it doubly-exponentially improves over \citet{Hazan-Kale-2011}'s regret on its
%dependence on the diameter of the benchmark class.

%\citet{Chen-Lin-Lu-2014}'s online weak learning condition implies that the set
%of examples is strongly separable by a convex combination of base hypotheses
%with a margin. Under this condition, it gives an algorithm with a $O(T^{3/4})$
%mistake bound. \citet{Zhang-Jung-Tewari-2018} considers an online weak learning
%condition, which implies that all examples is separable by a convex combination
%of base hypotheses with a margin; see~\citet[Theorem
%3]{Mukherjee-Schapire-2013}. Under this condition, it gives a boosting
%algorithm with a $O(\sqrt{T})$ mistake bound with the knowledge of the edge
%parameter. In addition, it gives an adaptive algorithm with a $O(T^{3/4})$
%mistake bound.

%Online bandit multiclass learning is a special case of the contextual bandits
%problem~\citep{Auer-2003, Langford-Zhang-2008}, where the cost is the
%classification error. A series of works in the contextual bandits literature
%focuses on oracle-efficiency, that is, they assume access to an oracle that
%given a set of cost-sensitve learning examples, generates .

There is a large body of literature on the contextual bandits
problem~\citep{Auer-2003, Langford-Zhang-2008}. In this problem, the learner is
given a policy class $\Pi$, and at each timestep $t$, an example $x_t$ is shown,
with an associated cost vector $c_t$ hidden. The learner then takes an action
$\widehat{y}_t$, and observes the incurred cost $c_t(\widehat{y}_t)$. The goal
of the learner is to minimize its regret $\sum_{t=1}^T c_t(\widehat{y}_t) -
\min_{\pi \in \Pi} \sum_{t=1}^T c_t(\pi(x_t))$. The online bandit multiclass
learning problem can be seen a special case of the contextual bandits problem, where the
cost $c_t(i)$ is the classification error $\indicator{i \neq y_t}$, and the
policy class $\Pi$ is the set of linear classifiers $\cbr{x \mapsto \argmax_y
(Wx)_y: W \in \R^{k \times d}}$. A series of works in the contextual bandits
learning literature focus on
oracle-efficiency~\citep{Dudik-Hsu-Kale-Karampatziakis-Langford-Reyzin-Zhang-2011,
Agarwal-Hsu-Kale-Langford-Li-Schapire-2014, Rakhlin-Sridharan-2016,
Syrgkanis-Krishnamurthy-Schapire-2016,
Syrgkanis-Luo-Krishnamurthy-Schapire-2016}. Specifically, they assume access to
a policy optimization oracle that can return the policy in $\Pi$ that minimizes
its empirical cost on any set of cost-sensitive examples, and the algorithm only
calls the oracle a polynomial number of times. However, these algorithms are not
truly computationally efficient in our setting, as it is NP-hard in general to
find a linear classifier that minimizes the empirical cost over a set of
examples~\citep{Arora-Babai-Stern-Sweedyk-1997}.

%that receives a set of cost-sensitive learning examples and policy class $\Pi$,
%as algorithms generate cost-sensitive examples that may not be linearly
%separable Algorithms that satisfy oracle efficiency has been proposed in the
%literature, for example.

Recently, \citet{Foster-Krishnamurthy-2018} developed a rich theory of
contextual bandits with surrogate losses, focusing on regrets of the form
$\sum_{t=1}^T c_t(\widehat{y}_t) - \min_{f \in \calF} \sum_{t=1}^T \frac{1}{K}
\sum_{i=1}^K c_t(i) \phi( f_i(x_t) )$, where $\calF$ contains score functions
$f$ such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac
s \gamma, 0)$ or $\min(1, \max(1 - \frac s \gamma, 0))$. On one hand, it gives
information-theoretic regret upper bounds under various settings of $\calF$.
On the other hand, it gives an
efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of
$\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$. A direct
application of this result to \textsc{Online Bandit Multiclass Linear
Classification} yields an algorithm with a $O(\sqrt{T})$ mistake bound in the
strongly linearly separable setting.

%including parametric and nonparametric classes

%\cite{Chen-Chen-Zhang-Chen-Zhang-2009} studies the approach of reducing bandit
%multiclass learning to online binary classification using one-versus-all
%reduction. They show that

%Their results do not imply a finite mistake bound in the weakly separable
%setting, in that the benchmark loss can still be $\Omega(T)$.