### 2019-01-22 ICML-submission/appendix-nearest-neighbor-algorithm.tex
Clean-ups

\section{Nearest neighbor algorithm}
\label{section:nearest-neighbor-algorithm}

\begin{algorithm}[H]
\caption{\textsc{Nearest-Neighbor Algorithm}}
\begin{algorithmic}[1]
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\REQUIRE{Inner product space $(V,\ip{\cdot}{\cdot})$.}
\STATE $S \gets \emptyset$.
\FOR{$t=1,2,\ldots,T$:}
\IF{$\min_{(x,y) \in S} \norm{x_t - x} \le \gamma$}
  \STATE{Find nearest neighbor \\ $(\widetilde{x}, \widetilde{y}) = \argmin_{(x,y) \in S} \norm{x_t - x}$}
	\STATE{Predict $\widehat{y}_t = \widetilde{y}$}
\ELSE
	\STATE{Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$}
	\STATE{Receive feedback $z_t = \indicator{\widehat{y}_t \neq y_t}$}
  \label{line:nearest-neighbor-explore}
	\IF{$z_t = 0$}
    \STATE $S \gets S \cup \cbr{(x_t, \widehat{y}_t)}$.
  \ENDIF
\ENDIF
\ENDFOR
\end{algorithmic}
\label{algorithm:nearest-neighbor}
\end{algorithm}

In this section we analyze \textsc{Nearest-Neighbor Algorithm} shown as
Algorithm~\ref{algorithm:nearest-neighbor}. The algorithm is based on the idea
that two examples that are close to each other must have the same label. Lemma
below formalizes this intuition.

\begin{lemma}
\label{lemma:nearest-neighbor}
Let $(V,\ip{\cdot}{\cdot})$ be a vector space, $K$ be a positive integer and let
$\gamma$ be a positive real number. Suppose $(x_1,y_1), (x_2,y_2), \dots, (x_T,
y_T) \in V \times \{1,2,\dots,K\}$ are labeled examples that are weakly linearly
separable with margin $\gamma$. For $i$, $j$ in $\cbr{1,2,\dots,T}$, if
$\norm{x_i - x_j}_2 \le \gamma$ then $y_i = y_j$.
\end{lemma}

\begin{proof}
Suppose for the sake on contradiction that $y_i \neq y_j$. By
Definition~\ref{definition:linear-separability}, there exists a set of $k$
vectors $(w_1, \ldots, w_k)$, such that
conditions~\eqref{equation:weak-linear-separability-1}
and~\eqref{equation:weak-linear-separability-2} are satisfied.

Specifically,
\begin{align*}
\ip{w_{y_i} - w_{y_j}}{x_i} & \ge \gamma \; , \\
\ip{w_{y_j} - w_{y_i}}{x_j} & \ge \gamma \; .
\end{align*}
This implies that
$$
\ip{w_{y_i} - w_{y_j}}{x_i - x_j} \ge 2\gamma \; .
$$

On the other hand,
$$
\ip{w_{y_i} - w_{y_j}}{x_i - x_j} \le \norm{w_{y_i} - w_{y_j}} \cdot \norm{x_i - x_j} < \sqrt{2} \gamma
$$
where the first inequality is from Cauchy-Schwartz inequality, the second
inequality is from that $\norm{w_{y_i} - w_{y_j}} \le \sqrt{2(\norm{w_{y_i}}^2 +
\norm{w_{y_j}}^2)} \leq \sqrt{2}$ and our assumption on $x_i$ and $x_j$.
Therefore, we reach a contradiction.
\end{proof}

\begin{theorem}
\label{theorem:margin_at_upper}
Let $K$ and $d$ be positive integers and let $\gamma$ be a positive real number.
Suppose $(x_1,y_1), \ldots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$ are
labeled examples that are weakly linearly separable with margin $\gamma$ and
satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$. Then, the expected
number of mistakes made by Algorithm~\ref{algorithm:nearest-neighbor} is at most
$$
(K-1) \left( \frac{R}{\gamma} \right)^d \; .
$$
\end{theorem}

\begin{proof}
Let $M$ be the number of mistakes made by the algorithm. Let $b_t$ be the
indicator that line~\ref{line:nearest-neighbor-explore} is executed at timestep
$t$, i.e. we fall into the ``else'' case. Note that if $b_t = 0$, then by
Lemma~\ref{lemma:nearest-neighbor}, the prediction $\widehat{y}_t$ must equal
$y_t$, i.e. $z_t = 0$. Therefore, $M = \sum_{t=1}^T z_t = \sum_{t=1}^T b_t z_t$.
Let $U = \sum_{t=1}^T b_t (1-z_t)$. Clearly, $|S| = U$. Since $S$ is a
$\gamma$-packing over $\cbr{x \in \R^d ~:~ \norm{x} \le R}$, the size of $S$ is
at most $(\frac R \gamma)^d$, which is also an upper bound on $U$.

Note that when $b_t = 1$, $\widehat{y}_t$ is chosen uniformly at random, we have
$$
\Exp[ z_t | b_t = 1] = \frac{K-1}{K} \; .
$$
Therefore,
$$
\Exp[M] = \Exp \left[ \sum_{t=1}^T b_t z_t \right] = \frac{K-1}{K} \Exp \left[ \sum_{t=1}^T b_t \right] \; .
$$
On the other hand,
$$
\Exp[U] = \Exp \left[ \sum_{t=1}^T b_t (1-z_t) \right] = \frac 1 K \Exp \left[\sum_{t=1}^T b_t \right] \; .
$$
Therefore,
$$
\Exp[M] = (K-1) \Exp[U] \leq (K-1) \left(\frac{R}{\gamma} \right)^d \; .
$$
\end{proof}