### 2019-01-23 ICML-submission/appendix-nearest-neighbor-algorithm.tex
Clean-ups

\section{Nearest neighbor algorithm}
\label{section:nearest-neighbor-algorithm}

\begin{algorithm}[H]
\SetAlgoLined
\LinesNumbered
\caption{\textsc{Nearest-Neighbor Algorithm}
\label{algorithm:nearest-neighbor}
}
\textbf{Require:} Number of classes $K$, number of rounds $T$. \\
\textbf{Require:} Inner product space $(V,\ip{\cdot}{\cdot})$. \\
\nl Initialize $S \gets \emptyset$ \\
\nl \For{$t=1,2,\ldots,T$:}{
\nl  \If{$\min_{(x,y) \in S} \norm{x_t - x} \le \gamma$}{
\nl    Find nearest neighbor \\
       \qquad \qquad $(\widetilde{x}, \widetilde{y}) = \argmin_{(x,y) \in S} \norm{x_t - x}$ \\
\nl	  Predict $\widehat{y}_t = \widetilde{y}$
  }
  \Else{
\nl	  Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$
      \label{line:nearest-neighbor-explore} \\
\nl	  Receive feedback $z_t = \indicator{\widehat{y}_t \neq y_t}$ \\
\nl	  \If{$z_t = 0$}{
\nl      $S \gets S \cup \cbr{(x_t, \widehat{y}_t)}$
    }
  }
}
\end{algorithm}

In this section we analyze \textsc{Nearest-Neighbor Algorithm} shown as
Algorithm~\ref{algorithm:nearest-neighbor}. The algorithm is based on obvious
the idea that under weak linear assumption two examples that are close to each
other must have the same label. Lemma below formalizes this intuition.

\begin{lemma}[Separation lemma]
\label{lemma:separation-lemma}
Let $(V,\ip{\cdot}{\cdot})$ be a vector space, $K$ be a positive integer and let
$\gamma$ be a positive real number. Suppose $(x_1,y_1), (x_2,y_2), \dots, (x_T,
y_T) \in V \times \{1,2,\dots,K\}$ are labeled examples that are weakly linearly
separable with margin $\gamma$. For $i$, $j$ in $\cbr{1,2,\dots,T}$, if
$\norm{x_i - x_j}_2 \le \gamma$ then $y_i = y_j$.
\end{lemma}

\begin{proof}
Suppose for the sake on contradiction that $y_i \neq y_j$. By
Definition~\ref{definition:linear-separability}, there exists a set of $k$
vectors $(w_1, \ldots, w_k)$, such that
conditions~\eqref{equation:weak-linear-separability-1}
and~\eqref{equation:weak-linear-separability-2} are satisfied.

Specifically,
\begin{align*}
\ip{w_{y_i} - w_{y_j}}{x_i} & \ge \gamma \; , \\
\ip{w_{y_j} - w_{y_i}}{x_j} & \ge \gamma \; .
\end{align*}
This implies that
$$
\ip{w_{y_i} - w_{y_j}}{x_i - x_j} \ge 2\gamma \; .
$$

On the other hand,
$$
\ip{w_{y_i} - w_{y_j}}{x_i - x_j} \le \norm{w_{y_i} - w_{y_j}} \cdot \norm{x_i - x_j} < \sqrt{2} \gamma
$$
where the first inequality is from Cauchy-Schwartz inequality, the second
inequality is from that $\norm{w_{y_i} - w_{y_j}} \le \sqrt{2(\norm{w_{y_i}}^2 +
\norm{w_{y_j}}^2)} \leq \sqrt{2}$ and our assumption on $x_i$ and $x_j$.
Therefore, we reach a contradiction.
\end{proof}

We also need to define several notions. A subset $S \subseteq \R^d$ is called a
$\gamma$-packing if for any $x,x' \in S$ such that $x \neq x'$ we have $\norm{x -
x'} > \gamma$. The following lemma is standard. Also recall that $B(x,R) = \{ x'
\in \R^d ~:~ \norm{x' - x} \le R \}$ denotes the closed ball of radius $R$
centered a point $x$.

\begin{lemma}[Size of $\gamma$-packing]
\label{lemma:size-of-packing}
Let $\gamma$ and $R$ be positive real numbers.
If $S \subseteq B(\zero,R) \subseteq \R^d$ is a $\gamma$-packing then
$$
|S| \le \left( \frac{2R}{\gamma} + 1 \right)^d \; .
$$
\end{lemma}

\begin{proof}
If $S$ is a $\gamma$-packing then $\{ B(x,\gamma/2) ~:~ x \in S \}$
is a collection of disjoint balls of radius $\gamma$ that fit into $B(\zero,R + \gamma/2)$.
Thus,
$$
|S| \cdot \Vol(B(\zero, \gamma/2)) \le \Vol(B(\zero,R + \gamma/2))
$$
Hence,
\begin{align*}
|S|
& \le \frac{\Vol(B(\zero,R + \gamma/2))}{\Vol(B(\zero, \gamma/2))} \\
& = \left( \frac{R + \gamma/2}{\gamma/2} \right)^d \\
& = \left( \frac{2R}{\gamma} + 1 \right)^d \; .
\end{align*}
\end{proof}

\begin{theorem}[Mistake upper bound for \textsc{Nearest-Neighbor Algorithm}]
\label{theorem:mistake-bound-for-nearest-neighbor-algorithm}
Let $K$ and $d$ be positive integers and let $\gamma,R$ be a positive real
numbers. Suppose $(x_1,y_1), \ldots, (x_T, y_T) \in
\R^d \times \{1,2,\dots,K\}$ are labeled examples that are weakly linearly
separable with margin $\gamma$ and satisfy $\norm{x_1}, \norm{x_2}, \dots,
\norm{x_T} \le R$. Then, the expected number of mistakes made by
Algorithm~\ref{algorithm:nearest-neighbor} is at most
$$
(K-1) \left( \frac{2R}{\gamma} + 1\right)^d \; .
$$
\end{theorem}

\begin{proof}
Let $M$ be the number of mistakes made by the algorithm. Let $b_t$ be the
indicator that line~\ref{line:nearest-neighbor-explore} is executed at timestep
$t$, i.e. we fall into the ``else'' case. Note that if $b_t = 0$, then by
Lemma~\ref{lemma:separation-lemma}, the prediction $\widehat{y}_t$ must equal
$y_t$, i.e. $z_t = 0$. Therefore, $M = \sum_{t=1}^T z_t = \sum_{t=1}^T b_t z_t$.
Let $U = \sum_{t=1}^T b_t (1-z_t)$. Clearly, $|S| = U$. Since $S \subseteq B(\zero, R)$
is a $\gamma$-packing, $U = |S| \le (\frac{2R}{\gamma} + 1)^d$.

Note that when $b_t = 1$, $\widehat{y}_t$ is chosen uniformly at random, we have
$$
\Exp[ z_t ~|~ b_t = 1] = \frac{K-1}{K} \; .
$$
Therefore,
$$
\Exp[M] = \Exp \left[ \sum_{t=1}^T b_t z_t \right] = \frac{K-1}{K} \Exp \left[ \sum_{t=1}^T b_t \right] \; .
$$
On the other hand,
$$
\Exp[U] = \Exp \left[ \sum_{t=1}^T b_t (1-z_t) \right] = \frac 1 K \Exp \left[\sum_{t=1}^T b_t \right] \; .
$$
Therefore,
$$
\Exp[M] = (K-1) \Exp[U] \leq (K-1) \left(\frac{2R}{\gamma} + 1 \right)^d \; .
$$
\end{proof}