### 2019-03-10 ICML-submission/rebuttal.txt
some comments on the theoretical parts in the rebuttal

We thank all reviewers for their insightful comments. We will make a pass,
and incorporate all the detailed comments in the final version of the paper.

We believe that extending the algorithms to handle the general
contextual bandit learning setting (where each example is associated
with a cost vector) requires nontrivial modifications to the algorithm;
a first step towards this question would be to establish
regret guarantees for when the examples shown is still multiclass, but
linearly nonseparable - we leave this for future work.

Regarding Appendix G, we agree that it is not related to the main story
 of the paper. Nevertheless, we think that this is a baseline that has not
 appeared in the literature and we include it for completeness.

We will add some discussions on achievable mistake bounds in the weakly
linearly separable case,
and its interaction with computational efficiency requirements.
It is known that in (Daniely and Herbertal, 2013)
that when 1/gamma^2 < d, any algorithm must make \Omega(K / \gamma^2) mistakes,
and when 1/gamma^2 > K^3 d, any algorithm must make \Omega( d K^2 ) mistakes.
In addition, (Daniely and Herbertal, 2013) gives a computationally inefficient
 algorithm that has a mistake upper bound of \tilde{O}( \min( K/\gamma^2, d K^2) );
 the algorithm involves a key step of computing (margin-based) Littlestone dimensions of subsets of multiclass linear classifiers, which is not known how to do efficiently.