### 2019-05-08 ICML-camera-ready/main-all.tex
change some figures

\documentclass{article}

\input{commands}

\begin{document}

\input{title-and-abstract}
\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}
\input{algorithm-for-strongly-linearly-separable-data}
\input{from-weak-separability-to-strong-separability}
\input{empirical-comparison}

\section*{Acknowledgments} We thank Francesco Orabona and Wen Sun
for helpful initial discussions, and thank Adam Klivans and Rocco Servedio for helpful
discussions on~\citep{Klivans-Servedio-2008} and
pointing out the reference~\citep{Klivans-Servedio-2004}. We also thank Dylan Foster, 
Akshay Krishnamurthy, and Haipeng Luo for providing a candidate solution to our problem.

\bibliography{biblio}
%\bibliographystyle{plainnat}
\bibliographystyle{icml2019}

% For actual submission remove everything below.
% Use appendices only in supplementray material.
\onecolumn
\appendix
\clearpage

\input{appendix-multiclass-perceptron}
\input{appendix-proofs-for-strongly-separable-examples}
\input{appendix-poly-norm-to-rkhs-norm}
\input{appendix-proofs-of-polynomial-approximation-theorems}
\input{appendix-margin-transformation}
\input{appendix-experiment}
\input{appendix-nearest-neighbor-algorithm}
\input{np-hardness-of-the-weak-labeling-problem}
\input{mistake-lower-bound-for-ignorant-algorithms}

\end{document}